Improving cropland maps through tight integration of human and machine intelligence

### 2021-01-28 inst/manuscript/manuscript.Rmd
Manuscript updates

Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms. Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly neural network-based "deep-learning" methods (cite). To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g. @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019]. Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g. @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012]. Visually interpreted training labels present two particular problems. The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019]. These errors are typically not accounted for in reported accuracy metrics, despite the fact that they can introduce substantial error into the resulting maps and subsequent analyses [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018]. The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland. Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial image sources that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018]. This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g. Landsat, Sentinel; @XiongNominal30mCropland2017) and often represent a different year. If a land change occurs in the interval between the two image sets (e.g. a new field was created), this can introduce error into the training data that is then passed on to the classifier. This source of error may be elevated in smallholder-dominated systems, where swidden practices are common (cite), or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].  
<!-- - Why cropland mapping is hard -->
<!--     - 3 problems -->
<!--         - The target itself and its spectral variability and the spectral variability of its background -->
<!--             - Spectral and spatial resolution of sensors -->
<!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->

@@

We describe here a cropland mapping system that addresses these requirements on the way to creating an object-based map of individual fields. The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017]. Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016]. Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3). The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016] and was re-engineered to run on high performance, cloud-based computing clusters. This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error. The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017]. Our framework automates and formalizes this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g. @XiongNominal30mCropland2017; others add]. Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017]. Finally, an unsupervised  segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, **resulting in a further sharpening of accuracy** and resulting in a final vector-based field boundary map.

@@

# Results
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
                                 package = "activemapper"))
km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
km2 <- format(km2, big.mark = ",", scientific = FALSE)
```
We developed maps for all of Ghana (`r km2` km$^2$). Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops. For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or replanted tree crops. We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.

@@

- Left unquantified is how much training on base maps matters
- Image quality:
    - why mean-weighting not median composite (artifacts) (maybe methods)
    - impact on accuracy
- Assessment of segmentation quality
- Next steps: 
    - Annually updated cropland maps
    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.

>>>>

Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms. Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly deep-learning methods (cite). To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g. @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019]. Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g. @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012]. Visually interpreted training labels present two particular problems. The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019]. These errors are typically not accounted for in reported accuracy metrics, despite the fact that they can introduce substantial error into the resulting maps and subsequent analyses [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018]. The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland. Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial image sources that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018]. This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g. Landsat, Sentinel; @XiongNominal30mCropland2017) and often represent a different year. If a land change occurs in the interval between the two image sets (e.g. a new field was created), this can introduce error into the training data that is then passed on to the classifier. This source of error may be elevated in smallholder-dominated systems, where swidden practices are common (cite), or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].  
<!-- - Why cropland mapping is hard -->
<!--     - 3 problems -->
<!--         - The target itself and its spectral variability and the spectral variability of its background -->
<!--             - Spectral and spatial resolution of sensors -->
<!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->

@@

We describe here a cropland mapping system that addresses these requirements on the way to delineating field boundaries. The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017]. Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016]. Their daily revisit capacity also is important for being able to develop cloud-free, sub-annual composites in cloudy regions.    Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3). The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016] and was re-engineered to run on high performance, cloud-based computing clusters. This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error. The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017]. Our framework automates and formalizes this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g. @XiongNominal30mCropland2017; others add]. Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017]. Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, **resulting in a further sharpening of accuracy** and resulting in a final vector-based field boundary map.

@@

# Results
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
                                 package = "activemapper"))
km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
km2 <- format(km2, big.mark = ",", scientific = FALSE)
```
We developed maps for all of Ghana (`r km2` km$^2$). Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops. For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or replanted tree crops. We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are often not visible within cocoa crops.

@@

- Left unquantified is how much training on base maps matters
- Image quality:
    - why mean-weighting not median composite (artifacts) (maybe methods)
    - impact on accuracy
- Assessment of segmentation quality
- Next steps: 
    - Annually updated cropland maps
    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
- Relevant literature: 
    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values). Semi-automated approach to generating reference data from imagery (based on phenology)
    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours. 
    - Feng et al (2020) importance of temporal information (backs up Debats et al).
      - temporal information makes labelling harder though

Improvements:    
- Replace Random Forest with deep learning model. Can still retain the active architecture, but instead use new sites to improve a pre-trained model
- Map along agroecozone boundaries, rather than random blocks
